---
title: "README"
author: "Me"
date: "15/03/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

Preparation: Downloading data and assigning to dataframes

```{r download and assign, echo=TRUE}
library(data.table, warn.conflicts = FALSE)

filename <- "DataCleaningPeerReviewAssignment.zip"

# Check for file
if (!file.exists(filename)){
        fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
        download.file(fileURL, filename, method="curl")
}  

# Check for folder
if (!file.exists("UCI HAR Dataset")) { 
        unzip(filename) 
}

#read and assign to dataframes
features <- read.table("UCI HAR Dataset/features.txt", col.names = c("n","functions"))
activities <- read.table("UCI HAR Dataset/activity_labels.txt", col.names = c("code", "activity"))
subject_test <- read.table("UCI HAR Dataset/test/subject_test.txt", col.names = "subject")
x_test <- read.table("UCI HAR Dataset/test/X_test.txt", col.names = features$functions)
y_test <- read.table("UCI HAR Dataset/test/y_test.txt", col.names = "code")
subject_train <- read.table("UCI HAR Dataset/train/subject_train.txt", col.names = "subject")
x_train <- read.table("UCI HAR Dataset/train/X_train.txt", col.names = features$functions)
y_train <- read.table("UCI HAR Dataset/train/y_train.txt", col.names = "code")


```

## Tidying Steps

run_analysis.R tidies the data according to these steps:

1. Merges the training and the test sets to create one data set.
2. Extracts only the measurements on the mean and standard deviation for each measurement.
3. Uses descriptive activity names to name the activities in the data set
4. Appropriately labels the data set with descriptive variable names.
5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

```{r run analysis, echo=TRUE}
#load dplyr
library(dplyr, warn.conflicts = FALSE)

#Step 1: Merges the training and the test sets to create one data set.
#merging X,Y,Subject for train and test datasets into mergeddata dataframe
X <- rbind(x_train, x_test)
Y <- rbind(y_train, y_test)
Subject <- rbind(subject_train, subject_test)
mergeddata <- cbind(Subject, Y, X)

#Step 2: Extracts only the measurements on the mean and standard deviation for each measurement.
#Using dplyr SELECT subject and activity code, and only the variables containing "mean" and "std"
dat <- mergeddata %>% select(subject, code, contains("mean"), contains("std"))

#Step 3: Uses descriptive activity names to name the activities in the data set.
#replace codes with activity names from activities file
dat$code <- activities[dat$code, 2]

#Step 4: Appropriately labels the data set with descriptive variable names.
#rename variables with descriptions using gsub() 
names(dat)[2] = "Activity"
names(dat)<-gsub("Acc", "Accelerometer", names(dat))
names(dat)<-gsub("Gyro", "Gyroscope", names(dat))
names(dat)<-gsub("BodyBody", "Body", names(dat))
names(dat)<-gsub("Mag", "Magnitude", names(dat))
names(dat)<-gsub("^t", "Time", names(dat))
names(dat)<-gsub("^f", "Frequency", names(dat))
names(dat)<-gsub("tBody", "TimeBody", names(dat))
names(dat)<-gsub("-mean()", "Mean", names(dat), ignore.case = TRUE)
names(dat)<-gsub("-std()", "STD", names(dat), ignore.case = TRUE)
names(dat)<-gsub("-freq()", "Frequency", names(dat), ignore.case = TRUE)
names(dat)<-gsub("angle", "Angle", names(dat))
names(dat)<-gsub("gravity", "Gravity", names(dat))


#Step 5: From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.
# chaining group by subject and activity followed by summarize_all to get the average per subject and activity group
tidy <- dat %>%
        group_by(subject, Activity) %>%
        summarize_all(list(~mean))
write.table(tidy, "tidydata.txt", row.name=FALSE)
```

